---
title: "characteristics of each stage"
author: "Ilhan Yoo"
date: "r format(Sys.Date())"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center", message=F, warning=F, dpi = 300, dev="tiff")
```

## library loading

```{r}
library(tidyverse)
library(gridExtra)
```

## load datasets
# BMR stage dataset
# dead subjects dataset
# King and MiToS stage dataset, subjectid의 이름은 bmr stage dataset과 같은 SubjectID로 변경 
# ALSFRS-R score dataset

```{r}
bmr <- read.csv("stage_bmr.csv")
surv <- read.csv("survival.csv")
king_mitos <- read.csv("stage_king_mitos.csv") 
alsfrs_r <- read.csv("alsfrs_rev.csv")
```

## merge dataset for analysis
# BMR stage에 대응하는 ALSFRS-R 확인위해 데이터셋 merge ALSFRS-R score dataset and BMR stage dataset
# King, MiToS stage에 대응하는 ALSFRS-R 확인위해 데이터셋 merge ALSFRS-R score dataset and King and MiToS stage dataset
```{r}
alsfrs_bmr <- bmr %>% left_join(alsfrs_r,by=c("SubjectID","feature_delta"))
alsfrs_king_mitos <- king_mitos %>% left_join(alsfrs_r,by=c("SubjectID","feature_delta"))
```

## BMR stage 7인 subject들의 ALSFRS-R score확인 
# BMR stage 7은 사망한 상태이므로 ALSFRS-R score가 있다면 잘못된 데이터이므로 filtering필요
```{r}
alsfrs_bmr %>% filter(bmr_stage==7) %>% group_by(ALSFRS_R_Total) %>% tally()
```

## BMR stage 7인 subject의 ALSFRS-R 9,12,17,34인 경우 1명씩 제외, NA값은 남겨놓음 

```{r}
temp <- alsfrs_bmr %>% filter(bmr_stage==7&!is.na(ALSFRS_R_Total))
alsfrs_bmr1 <- alsfrs_bmr %>% anti_join(temp)
rm(temp)
```
## ALSFRS total score and subscore, which were divided into 3 domains from ALSFRS total score, according to BMR stage
```{r}
alsfrs_bmr1 %>% group_by(bmr_stage) %>% summarise(alsfrs_mean=mean(ALSFRS_R_Total),alsfrs_sd=sd(ALSFRS_R_Total),
                                                 bulbar_mean=mean(bulbar),bulbar_sd=sd(bulbar),
                                                 motor_mean=mean(motor),motor_sd=sd(motor),
                                                 respiratory_mean=mean(respiratory),respiratory_sd=sd(respiratory))
```
## 사망하여 ALSFRS-R이 존재할 수 없는 stage 7를 제외한 나머지 BMR stage와 ALSFRS-R association 확인하는 plot  
```{r}
p_bmr_1 <- alsfrs_bmr1 %>% ggplot(aes(x=factor(bmr_stage),y=ALSFRS_R_Total))+
  geom_boxplot(width=0.8,outlier.size=3,outlier.shape=16,outlier.color = "red")+
  theme(legend.position = "none",panel.grid.major = element_blank(),panel.grid.minor = element_blank())+
  theme_classic()+
  labs(x="BMR stage",y="ALSFRS-R")
```

## King's stage 5인 subject들의 ALSFRS-R score확인 
# King's stage 5는 사망한 상태이므로 ALSFRS-R score가 있다면 잘못된 데이터이므로 filtering필요
```{r}
alsfrs_king_mitos %>% filter(king==5) %>% group_by(ALSFRS_R_Total) %>% tally()
```

## King's stage 5인 subject의 ALSFRS-R 9,12,17,34인 경우 1명씩 제외, NA값은 남겨놓음 
```{r}
temp <- alsfrs_king_mitos %>% filter(king==5&!is.na(ALSFRS_R_Total))
alsfrs_king_mitos1 <- alsfrs_king_mitos %>% anti_join(temp)
rm(temp)
```

## ALSFRS total score and subscore, which were divided into 3 domains from ALSFRS total score, according to King stage
```{r}
alsfrs_king_mitos %>% group_by(king) %>% summarise(alsfrs_mean=mean(ALSFRS_R_Total),alsfrs_sd=sd(ALSFRS_R_Total),
                                                 bulbar_mean=mean(bulbar),bulbar_sd=sd(bulbar),
                                                 motor_mean=mean(motor),motor_sd=sd(motor),
                                                 respiratory_mean=mean(respiratory),respiratory_sd=sd(respiratory))
```

## King's stage-ALSFRS-R plot without King's stage 5
#사망하여 ALSFRS-R이 존재할 수 없는 stage 5를 제외한 나머지 King's stage와 ALSFRS-R association 확인하는 plot
```{r}
p_king_1 <- alsfrs_king_mitos1 %>% ggplot(aes(x=factor(king),y=ALSFRS_R_Total))+
  geom_boxplot(width=0.8,outlier.size=3,outlier.shape=16,outlier.color = "red")+
  theme(legend.position = "none",panel.grid.major = element_blank(),panel.grid.minor = element_blank())+
  theme_classic()+
  coord_cartesian(xlim = c(1,4))+
  labs(x="King's stage",y="ALSFRS-R")
```

## MiToS stage 5인 subject들의 ALSFRS-R score확인 
# MiToS stage 5는 사망한 상태이므로 ALSFRS-R score가 있다면 잘못된 데이터이므로 filtering필요
```{r}
alsfrs_king_mitos %>% filter(mitos==5) %>% group_by(ALSFRS_R_Total) %>% tally()
```

## MiToS stage 5인 subject의 ALSFRS-R 9,12,17,34인 경우 1명씩 제외, NA값은 남겨놓음 
```{r}
temp <- alsfrs_king_mitos %>% filter(king==5&!is.na(ALSFRS_R_Total))
alsfrs_king_mitos1 <- alsfrs_king_mitos %>% anti_join(temp)
rm(temp)
```

## ALSFRS total score and subscore, which were divided into 3 domains from ALSFRS total score, according to MiToS stage
```{r}
alsfrs_king_mitos1 %>% group_by(mitos) %>% summarise(alsfrs_mean=mean(ALSFRS_R_Total),alsfrs_sd=sd(ALSFRS_R_Total),
                                                 bulbar_mean=mean(bulbar),bulbar_sd=sd(bulbar),
                                                 motor_mean=mean(motor),motor_sd=sd(motor),
                                                 respiratory_mean=mean(respiratory),respiratory_sd=sd(respiratory))
```

## MiToS stage-ALSFRS-R plot without MiToS stage 5
#사망하여 ALSFRS-R이 존재할 수 없는 stage 5를 제외한 나머지 MiToS stage와 ALSFRS-R association 확인하는 plot 
```{r}
p_mitos_1 <- alsfrs_king_mitos1 %>% ggplot(aes(x=factor(mitos),y=ALSFRS_R_Total))+
  geom_boxplot(width=0.8,outlier.size=3,outlier.shape=16,outlier.color = "red")+
  theme(legend.position = "none",panel.grid.major = element_blank(),panel.grid.minor = element_blank())+
  theme_classic()+
  coord_cartesian(xlim = c(1,6))+
  labs(x="MiToS stage",y="ALSFRS-R")
```

# BMR stage system이 ALSFRS-R의 변화를 잘 반영하는지를 King's,MiToS stage과 비교하기위해
# BMR,King,MiToS stage plot merge시킴
```{r}
grid.arrange(p_bmr_1,p_king_1,p_mitos_1,ncol=1)
```